Left-linear rewriting in adhesive categories

### 2024-04-21 concur/fusione.tex
typo intro

A noteworthy advantage is that DPO rewriting can be formulated over
adhesive categories and their variants~\cite{lack2005adhesive,ehrig2006weak}. 
%
Operating within the framework of adhesive categories allows one to
devise the foundations of a rewriting theory that can be subsequently
instantiated to the context of interest, thus avoiding the need of
repeatedly proving similar ad-hoc results for each specific
setting. In fact, the framework is so general that it allowed to
recast in the DPO approach many additional formalisms, ranging from
toposes~\cite{johnstone2007quasitoposes} to string
diagrams~\cite{bonchi2022string}.

@@

As an initial step, % in performing this exploration, 
we observe that for
left-linear rules a number of fundamental properties that play a role
in the concurrency theory of linear rewriting need to be deeply revised, or
plainly do not hold.
%
Firstly, the notion of sequential independence and the Church-Rosser
theorem, ensuring that sequential independent steps can be performed
in reverse order, does not hold out of the box. Even worse, given two
independent steps it might be possible to switch them in several
different ways.
%
As we will observe, this prevents the development of a sensible theory
of concurrency since switches are used as a basis for equating
sequences of rewriting steps that differs only for the order of
independent steps.

@@

On the one hand, they capture most categories of
interest for rewriting. In particular, we identify a vast class of
left-linear rewriting systems on graph-like structures that are
well-switching and we show that rewriting systems on graphs with
equivalences are so. 
As a sanity check, we prove that linear rewriting systems on
$\mathcal{M}$-adhesive categories are well-switching.

>>>>

A noteworthy advantage of DPO rewriting is that it can be formulated over
adhesive categories and their variants~\cite{lack2005adhesive,ehrig2006weak}. 
%
Operating within the framework of adhesive categories allows one to
devise the foundations of a rewriting theory that can be subsequently
instantiated to the context of interest, thus avoiding the need of
repeatedly proving similar ad-hoc results for each specific
setting. In fact, the framework is so general that it allowed to
recast in the DPO approach many additional formalisms, ranging from
toposes~\cite{johnstone2007quasitoposes} to string
diagrams~\cite{bonchi2022string}.

@@

As an initial step, % in performing this exploration, 
we observe that for
left-linear rules a number of fundamental properties that play a role
in the concurrency theory of linear rewriting need to be deeply revised, or
plainly do not hold.
%
Firstly, the notion of sequential independence and the Church-Rosser
theorem, ensuring that sequential independent steps can be performed
in reverse order, does not hold out of the box. Even worse, given two
independent steps it might be possible to switch them in several
different ways.
%
As we will observe, this prevents the development of a sensible theory
of concurrency, since switches are used as a basis for equating
sequences of rewriting steps that differ only for the order of
independent steps.

@@

On the one hand, they capture most categories of
interest for rewriting. In particular, we identify a vast class of
left-linear rewriting systems on graph-like structures that are
well-switching and we show that rewriting systems on graphs with
equivalences are so.\todo{weaker statement?}
As a sanity check, we prove that linear rewriting systems on
$\mathcal{M}$-adhesive categories are well-switching.
